---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02

I have selected for this project the dataset about bad drivers in U.S. I pretend to visualize which State has the worst drivers in the country and if the number of accidents where drivers who were not distracted is considerable. I am not the best driver but I drive without exceeding the limit of speed, and I am always afraid of crashes because Floridian's drivers are really racing and that scares me a little bit. 


## Getting the data



```{r}
library(tidyverse)
bad_drivers<-read_csv ("https://raw.githubusercontent.com/reisanar/datasets/master/bad_drivers.csv")

```

This dataset contains the following variables:

* State
* Number of drivers involved in fatal collisions per billion miles
* Percentage of drivers involved in fatal collisions who were speeding
* Percentage of drivers involved in fatal collisions who were alcohol-impaired
* Percentage of drivers involved in fatal collisions who were not distracted
* Percentage of drivers involved in fatal collisions who had not been involved in any previous accidents
* Car insurance premiums
* Losses incurred by insurance companies for collisions per insured driver



# Summarizing the data

After getting the data, it is time to formatting the data because the columns "perc_speeding", "perc_alcohol", "perc_not_distracted" and so on, are percentages of "num_drivers (number of drivers involved in fatal accidents)". I will mutate some new columns in the following step:



```{r}
bad_drivers<-bad_drivers%>%
  mutate(speeding_drivers=(num_drivers*perc_speeding)/100)%>%
  mutate(alchol_drivers=(num_drivers*perc_alcohol)/100)%>%
  mutate(notdistracted_drivers=(num_drivers*perc_not_distracted)/100)%>%
  mutate(noprevious_drivers=(num_drivers*perc_no_previous)/100)

```





# Plotting the data

Below will be a Treemap that will show us the number of bad drivers in U.S grouping by State.



```{r}
library(ggplot2)
library(treemapify)

treemap1<- ggplot(bad_drivers, aes(area = num_drivers, fill = state, label = paste(state, num_drivers, separate = " "))) + 
  geom_treemap(show.legend = FALSE) +
  geom_treemap_text(fontface = "italic", colour = "white", place = "topleft", grow = FALSE) +
  labs(title = "Number of Bad Drivers in U.S", subtitle = "Per State")

treemap1

```


